\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-28 paper/main.tex
minor update BDM-DG

\subsection{Discretization of the Stokes equations}
We outline the discretization of the weak
formulation of the Stokes equations for the baseline model 
(both cilia traction and cardiac pressure pulsations present) given by
\Cref{eq:abstract_weak_form_modelC,eq:stokes_bilinear_form,eq:stokes_skew_form,eq:stokes_linear_form}.
A similar approach can be used to discretize the weak formulation
for the other model versions.
First, we consider the spatial discretization of the weak formulation.
We rewrite the weak form in terms of the discretized variables,
denoted with the subscript $h$, on the discrete domain $\Omega_h$:
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h)
        &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:no_stabilization_abstract_discrete_weak_form_modelC}%
\end{subequations}%
Our choice of discretization are finite elements
of the Brezzi-Douglas-Marini (BDM) family~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
for the velocity
space and discontinuous Lagrange polynomial elements for the pressure. We refer to the latter
as discontinuous Galerkin (DG) elements. For all cells $K$ of the computational mesh $\Omega_h$,
we define
\begin{align*}
    \VV_h &= \Big\{\vv_h : \vv_h\in\mathbf{H}(\mathrm{div});
    \vv_h\vert_K\in\big[P_k(K)\big]^d;
    \forall K\in\Omega_h;
    \vv_h\cdot\nn=0 \ \text{on } \Gamma\setminus\Gp\Big\},\\
    Q_h &= \Big\{q_h : q_h\in L^2;  q_h\vert_K\in P_{k-1}(K); \forall K\in\Omega_h\Big\},
\end{align*}
where $P_k$ is the space of polynomials of degree less than or equal to $k$.
The BDM family of elements have degrees of freedom
associated with integral moments of vector components normal to facets of the mesh.
Because $\VV_h\not\subset\VV$, the elements are non-conforming.
This BDM-DG scheme for the Stokes equations
is $\mathbf{H}(\mathrm{div})$ conforming and thus exactly mass-conserving,
because the continuity equation can be satisfied
in every computational mesh cell due to the fact that $\nabla\cdot\VV_h\subset Q_h$.
Moreover, control of normal components of $\uu_h$ conveniently 
lets us enforce the impermeability condition $\uu_h\cdot\nn=0$ strongly.
requires a stabilization of the bilinear form $\mathcal{A}(\uu_h, \vv_h)$.

Because the elements of $\VV_h$ define the normal components of $\uu_h$,
continuity in tangential components of $\uu_h$ on interior facets of
the mesh is ensured through stabilization
of the discrete weak form defined by \Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}.
We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}.
First, we define the average and jump operators, respectively, as
\begin{equation}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2}
    \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
    \label{eq:operators_BDM}
\end{equation}
where $\uu$ is a first-order tensor and 
$\nn$ is the normal vector on a facet separating two cells,
and the plus and minus signs denote variables in each of the two cells.
The normal vector is chosen such that it points towards the exterior of
its corresponding cell. On the boundary, the operators are defined as
$\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. Note that average operator
defined in~\Cref{eq:operators_BDM} reduces the rank of the tensor $\uu$ by one,
whereas the jump operator preserves rank. This extends to tensors of higher order.
We now define and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &=
    \sum_{f\in \mathcal{F}}\int_f{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}
    \cdot\jump{P_{\nn}(\vv_h)}}\,\mathrm ds\\
    &-\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}\,\mathrm ds
    -\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}\,\mathrm ds
\end{align*}
to the discrete weak formulation (\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}),
where $\mathcal{F}$ is the set of all interior facets of the computational mesh.
The first term of $\mathcal{A}_{\mathrm{stab}}$ penalizes jumps
in the tangential components of the velocity
across interior facets, where the penalty parameter $\gamma>0$ must be chosen large enough
to achieve stability. For appropriate convergence, the penalization term
includes a scaling with the average of the cell diameter of neighboring mesh cells,
with $h$ defined as the maximal distance between two points in a cell~\cite{Nitsche1971UberSind}.
The other two terms
of $\mathcal{A}_{\mathrm{stab}}$ are added to ensure consistency and symmetry of
the bilinear form, such that the problem is well-defined.

>>>>

\subsection{Discretization of the Stokes equations}
We outline the discretization of the weak
formulation of the Stokes equations for the baseline model 
(both cilia traction and cardiac pressure pulsations present) given by
\Cref{eq:abstract_weak_form_modelC,eq:stokes_bilinear_form,eq:stokes_skew_form,eq:stokes_linear_form}.
A similar approach can be used to discretize the weak formulation
for the other model versions.
First, we consider the spatial discretization of the weak formulation.
We rewrite the weak form in terms of the discretized variables,
denoted with the subscript $h$, on the discrete domain $\Omega_h$:
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h)
        &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:no_stabilization_abstract_discrete_weak_form_modelC}%
\end{subequations}%
Our choice of discretization are finite elements
of the Brezzi-Douglas-Marini (BDM) family~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
for the velocity
space and discontinuous Lagrange polynomial elements for the pressure. We refer to the latter
as discontinuous Galerkin (DG) elements. For all cells $K$ of the computational mesh $\Omega_h$,
we define
\begin{align*}
    \VV_h &= \Big\{\vv_h : \vv_h\in\mathbf{H}(\mathrm{div});
    \vv_h\vert_K\in\big[P_k(K)\big]^d;
    \forall K\in\Omega_h;
    \vv_h\cdot\nn=0 \ \text{on } \Gamma\setminus\Gp\Big\},\\
    Q_h &= \Big\{q_h : q_h\in L^2;  q_h\vert_K\in P_{k-1}(K); \forall K\in\Omega_h\Big\},
\end{align*}
where $P_k$ is the space of polynomials of degree less than or equal to $k$.
The BDM family of elements have degrees of freedom
associated with integral moments of vector components normal to facets of the mesh.
Because $\VV_h\not\subset\VV$, this BDM-DG scheme yields a non-conforming
method for the Stokes equations. The scheme is
$\mathbf{H}(\mathrm{div})$-conforming and exactly mass-conserving:
due to the fact that $\nabla\cdot\VV_h\subset Q_h$,
the continuity equation (\Cref{eq:BDM_stokes_div})
is satisfied in every computational mesh cell.
Moreover, control of normal components of $\uu_h$ conveniently 
lets us enforce the impermeability condition $\uu_h\cdot\nn=0$ strongly.

Because the elements of $\VV_h$ only define the normal components of $\uu_h$,
continuity in tangential components of $\uu_h$ on interior facets of
the mesh is ensured through stabilization
of the discrete weak form (\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}).
We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}.
First, we define the average and jump operators, respectively, as
\begin{equation}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2}
    \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
    \label{eq:operators_BDM}
\end{equation}
where $\uu$ is a first-order tensor and 
$\nn$ is the normal vector on a facet separating two cells,
and the plus and minus signs denote variables in each of the two cells.
The normal vector is chosen such that it points towards the exterior of
its corresponding cell. On the boundary, the operators are defined as
$\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. Note that average operator
defined in~\Cref{eq:operators_BDM} reduces the rank of the tensor $\uu$ by one,
whereas the jump operator preserves rank. This extends to tensors of higher order.
We now define and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &=
    \sum_{f\in \mathcal{F}}\int_f{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}
    \cdot\jump{P_{\nn}(\vv_h)}}\,\mathrm ds\\
    &-\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}\,\mathrm ds
    -\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}\,\mathrm ds
\end{align*}
to the discrete weak formulation (\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}),
where $\mathcal{F}$ is the set of all interior facets of the computational mesh.
The first term of $\mathcal{A}_{\mathrm{stab}}$ penalizes jumps
in the tangential components of the velocity
across interior facets, where the penalty parameter $\gamma>0$ must be chosen large enough
to achieve stability. For appropriate convergence, the penalization term
includes a scaling with the average of the cell diameter of neighboring mesh cells,
with $h$ defined as the maximal distance between two points in a cell~\cite{Nitsche1971UberSind}.
The other two terms
of $\mathcal{A}_{\mathrm{stab}}$ are added to ensure consistency and symmetry of
the bilinear form, such that the problem is well-defined.
